---
title: "Multiple phi, second example teething issues."
author: "Andrew Manderson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
fontfamily: tgpagella
fontsize: 10pt
papersize: a4
geometry: margin=2.25cm
bibliography: ../bibliography/multi-phi-bib.bib
csl: ../bibliography/journal-of-the-royal-statistical-society.csl
output: 
  html_document:
    code_folding: hide
---

```{r setup, include = FALSE, cache = FALSE}
options(width = 9999)
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```


I am currently trying to get the survival example to work in the multiple $\phi$ setting.
I can fit the uncertain event time submodel $\text{p}_{1}(\phi_{1 \cap 2}, \psi_{1}, Y_{1})$.
The next step is to meld it with a standard Weibull survival model.

# preliminary tests

As a preliminary test, I take one sample of the event times from $\text{p}_{1}$'s posterior, and treating it as data in a standard survival model. 
I also simulate one baseline covariate, with non-zero effect.
This lets me compare the results from my implementation to standard survival model implementations. 
To make the outputs roughly comparable, I have to centre the covariate and scale the event time to $[0, 1]$.
Even then, I can't get the various survival packages to agree on the sign
of the coefficients nor the estimate of the hazard shape $\lambda$.

```{r pkgs, include = FALSE, message = FALSE, warning = FALSE, comment = NA}
library(rstan) 
library(rstanarm) # for stan_surv
library(flexsurv) # for flexsurvreg
library(bayesplot)

source("../scripts/surv-example/GLOBALS.R")

set.seed(sim_seed)
submodel_one_output <- readRDS("../rds/surv-example/submodel-one-output.rds")
submodel_two_data <- readRDS("../rds/surv-example/submodel-two-simulated-data.rds")
psi_step_stan_model <- stan_model("../scripts/surv-example/models/submodel-two-psi-step.stan")
event_time_names <- sprintf("event_time[%d]", 1 : n_patients)

stan_data <- list(
  n_patients = length(submodel_two_data$patient_id),
  baseline_measurement = submodel_two_data$baseline_val, 
  event_time = submodel_one_output$samples[
    sample(1 : dim(submodel_one_output$samples)[1], 1), 
    sample(1 : dim(submodel_one_output$samples)[2], 1), 
    event_time_names
  ]
)

rescaled_stan_data <- stan_data
rescaled_stan_data$event_time <- with(stan_data, event_time / max(event_time))
rescaled_stan_data$baseline_measurement <- scale(
    stan_data$baseline_measurement,
    center = TRUE,
    scale = FALSE
  ) %>% 
  as.numeric()

surv_reg_df <- data.frame(
  event_time = rescaled_stan_data$event_time,
  event_indicator = rep(1, rescaled_stan_data$n_patients),
  baseline = rescaled_stan_data$baseline_measurement
)

```


# My output

```{r my_implementation, message = FALSE, warning = FALSE, comment = NA}
test_fit <- sampling(
  psi_step_stan_model,
  data = rescaled_stan_data,
  cores = 4,
  refresh = 0
)

print(test_fit, digits = 4, probs = c(0.025, 0.975))
```
Which uses the following Stan model/code, which I think is right (it is based of Appendix A.2 of @brilleman_bayesian_2020):

<details><summary>Show code</summary>
```{r my_stan_code, echo = FALSE, comment = NA}
cat(readLines("../scripts/surv-example/models/submodel-two-psi-step.stan"), sep = "\n")
```
</details>

# stan_surv

This uses Stan to sample a Weibull survival model as defined in @brilleman_bayesian_2020.

```{r stan_surv, message = FALSE, warning = FALSE, comment = NA}
res_surv_stan <- stan_surv(
  Surv(event_time, event_indicator) ~ baseline, 
  data = surv_reg_df, 
  basehaz = "weibull",
  cores = 4,
  refresh = 0
)

print(res_surv_stan, digits = 4)
```


# flexsurv

`flexsurv` doesn't seem to report its estimate of the intercept, and has two parameters for the Weibull distribution, shape and scale.
The versions of the Weibull survival model I've read all only have one (shape), maybe other packages do some kind of transformation of the response? 

```{r flexsurv, message = FALSE, warning = FALSE, comment = NA}
res_flex <- flexsurvreg(
  Surv(event_time, event_indicator) ~ 1 + baseline, 
  data = surv_reg_df, 
  dist = "weibull",
)

print(res_flex, digits = 4)
```

# Bibliography
